crop

### 2023-05-09 scripts/crop2boundary.Rmd
battling plss

```

```{r import states to crop to}
counties <- st_read('../geodata/USCounties')
st <- tigris::states() %>% select(STATEFP, STUSPS, STATE = NAME) %>% 
  filter(STUSPS %in% c('AZ', 'AK', 'CA', 'CO', 'ID', 'KS', 'NE', 'ND', 'NM', 'NV', 
                       'MN', 'MT', 'SD', 'TX', 'OK', 'OR', 'UT', 'WA', 'WY'))
```

```{r}
ggplot() + 
  geom_sf(data = st, aes(fill = STATEFP), color = NA)
```

>>>>

rm(mason)
```

```{r crop political boundary data to extent, eval = F}
counties <- st_read('../geodata_raw/USCounties/tl_2020_us_county.shp', quiet = T) %>% 
  st_transform(st_crs(tile_cells)) %>% 
  select(STATEFP, COUNTY = NAME)

states <- tigris::states() %>% 
  select(STATEFP, STATE = NAME, STUSPS) %>% 
  st_drop_geometry()

counties <- counties[st_intersects(counties, st_union(tile_cells)) %>% lengths > 0, ]
counties <- left_join(counties, states, by = "STATEFP") %>% 
  select(-STATEFP) %>% 
  arrange(STATE)

states <- counties %>%
  distinct(STUSPS) %>% 
  pull(STUSPS)

st_write(counties, dsn = file.path('../geodata/political', 'political.shp'))

rm(counties)
```

```{r crop mountain names database to extent, eval = F}

mtns <- st_read('../geodata_raw/globalMountains/GMBA_Inventory_v2.0_standard_basic.shp', 
        quiet = T) %>% 
  st_make_valid() %>% 
  select(MapName) %>% 
  st_crop(., st_union(tile_cells))

st_write(mtns, dsn = file.path('../geodata/mountains', 'mountains.shp'), quiet = T)

rm(mtns)
```

```{r crop geographic place name database to extent, eval = F}

files <- paste0('../geodata_raw/GNIS/Text/DomesticNames_', states, '.txt')
cols <- c('feature_name', 'prim_lat_dec', 'prim_long_dec')

places <- lapply(files, read.csv, sep = "|") %>% 
  map(., ~ .x |> select(all_of(cols))) %>% 
  data.table::rbindlist() %>% 
  st_as_sf(coords = c('prim_long_dec', 'prim_lat_dec'), crs = 4269) %>% 
  st_transform(4326)

places <- places[st_intersects(places, st_union(tile_cells)) %>% lengths > 0, ]

st_write(places, dsn = file.path('../geodata/places', 'places.shp'), quiet = T)

rm(places, cols, files)
```

```{r crop padus to domain, eval = F}

padus <- st_read(dsn = '../geodata_raw/PADUS3/PAD_US3_0.gdb', 
                 layer = 'PADUS3_0Fee') %>% 
  filter(State_Nm %in% states) %>% 
  select(Mang_Name, Unit_Nm) %>% 
  st_cast('MULTIPOLYGON')

tile_cells <- st_transform(tile_cells, st_crs(padus))
padus <- padus[st_intersects(padus, st_union(tile_cells)) %>% lengths > 0, ]
padus <- st_transform(padus, crs = 4326)

st_write(padus, dsn = file.path('../geodata/pad', 'pad.shp'), quiet = T)
rm(padus)
```

```{r geological map, eval = F}
geological <- st_read(
  '../geodata_raw/geologicalMap/USGS_StateGeologicMapCompilation_ver1.1.gdb', 
          layer = 'SGMC_Geology', quiet = T) %>% 
  select(GENERALIZED_LITH, UNIT_NAME)

tile_cells <- st_transform(tile_cells, st_crs(geological))
geological <- geological[st_intersects(geological, st_union(tile_cells)) %>% lengths > 0, ]
geological <- st_crop(geological, st_union(tile_cells))
geological <- st_transform(geological, 4326)

st_write(geological, dsn = file.path('../geodata/geology', 'geology.shp'), quiet = T)

tile_cells <- st_transform(tile_cells, crs = 4326)
rm(geological)
```

```{r allotments, eval = F}

blm <- st_read(
  '../geodata_raw/BLMAllotments/BLM_Natl_Grazing_Allotment_Polygons.shp') %>% 
  select(ALLOT_NAME)
usfs <- st_read(
  '../geodata_raw/USFSAllotment/S_USA.Allotment.shp', quiet = T
) %>% 
  select(ALLOT_NAME = ALLOTMENT_)

allotments <- bind_rows(blm, usfs) %>% 
  st_make_valid()
tile_cells <- st_transform(tile_cells, st_crs(allotments))
allotments <- allotments[st_intersects(allotments, st_union(tile_cells)) %>% lengths > 0, ]
allotments <- st_crop(allotments, st_union(tile_cells))
allotments <- st_transform(allotments, 4326)

st_write(allotments, dsn = file.path('../geodata/allotments', 'allotments.shp'), quiet = T)

rm(blm, usfs, allotments)
```

```{r public land survey system only for west the front range}

bound_nm <- data.frame(
  y = c(31, 31, 50, 50, 31), 
  x = c(-105, -125, -125, -105, -103)
) %>% 
  st_as_sf(coords = c('x', 'y'), crs = 4326) %>% 
  st_bbox() %>% 
  st_as_sfc()

township <- st_read(
  '../geodata_raw/nationalPLSS/ilmocplss.gdb', layer = 'PLSSTownship', quiet = T
) %>% 
  st_cast('POLYGON') %>% 
  select(TWNSHPLAB)

bound_nm <- st_transform(bound_nm, st_crs(township))
township <- township[st_intersects(township, bound_nm) %>% lengths > 0, ]

section <- st_read(
  '../geodata_raw/nationalPLSS/ilmocplss.gdb', layer = 'PLSSFirstDivision', quiet = T
) %>% 
  st_cast('POLYGON') %>% 
  select(FRSTDIVLAB)

section <- section[st_intersects(section, bound_nm) %>% lengths > 0, ]

trs <- st_join(township, section, .predicate = st_overlaps) %>% 
  mutate(TRS = paste0(TWNSHPLAB, ' S', FRSTDIVLAB)) %>% 
  select(TRS)

plot(trs)

before <- object.size(trs)
trs <- st_simplify(trs)
after <- object.size(trs)
after

st_write(trs, 
         dsn = file.path('../geodata/plss', 'plss.shp'), quiet = T)

rm(section, township, trs)
```
